---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, warning=FALSE}
library(dplyr)
library(ggplot2)
unzip("activity.zip")
activityData<-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
```{r, warning=FALSE}
activityByDay<-group_by(select(activityData,date,steps),date)
stepsPerDay<-summarise(activityByDay,sum(steps))

names(stepsPerDay)[2]<-"total.steps"

# hist(stepsPerDay$total.steps,xlab="Number of steps per day",
#      main="Histogram of number of steps taken per day",breaks=10,
#      col=topo.colors(1))
# 
# g<-ggplot(total.steps,aes(stepsPerDay))
# g<-g+geom_histogram()

g<-ggplot(stepsPerDay,aes(total.steps))
g<-g+geom_histogram(binwidth = 2500,col="white")
g<-g+ggtitle("Steps taken per day")
g<-g+labs(y="Frequency",x = "Total number of steps per day")
print(g)

meanStepsPerDay<-mean(stepsPerDay$total.steps,na.rm=TRUE)
medianStepsPerDay<-median(stepsPerDay$total.steps,na.rm=TRUE)


```

* Mean number of steps taken per day - ```r meanStepsPerDay ```
* Median number of steps per day - ```r medianStepsPerDay```


## What is the average daily activity pattern?

```{r}

stepsByTime<-group_by(select(activityData,interval,steps),interval)
stepsByTime<-summarise(stepsByTime,mean(steps,na.rm=TRUE))

names(stepsByTime)<-c("Interval","Average.Steps")

timeMin<-stepsByTime$Interval%%100
timeHour<-floor(stepsByTime$Interval/100)
stepsByTime$timeOfDay<-paste(timeHour,":",timeMin)

timeAtmax<-stepsByTime[stepsByTime$Average.Steps==max(stepsByTime$Average.Steps),]

g<-ggplot(stepsByTime,aes(y=Average.Steps,x=Interval,color=topo.colors(1)))
g<-g+geom_line()
g<-g+ theme(legend.position="none")
g<-g+labs(y="Average Steps",x = "Time in a day (5-min interval)")
print(g)

```

At 5-minute interval, on average across all the days maximum number of steps 
reported at ```r timeAtmax$timeOfDay```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
