---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---


## Loading and preprocessing the data
```{r, warning=FALSE}
library(dplyr)
library(ggplot2)
unzip("activity.zip")
activityData<-read.csv("activity.csv")
str(activityData)
```


## What is mean total number of steps taken per day?
```{r, warning=FALSE}
activityByDay<-group_by(select(activityData,date,steps),date)
stepsPerDay<-summarise(activityByDay,sum(steps))

names(stepsPerDay)[2]<-"total.steps"

g<-ggplot(stepsPerDay,aes(total.steps))
g<-g+geom_histogram(binwidth = 2500,col="white")
g<-g+ggtitle("Steps taken per day")
g<-g+labs(y="Frequency",x = "Total number of steps per day")
print(g)

meanStepsPerDay<-mean(stepsPerDay$total.steps,na.rm=TRUE)
medianStepsPerDay<-median(stepsPerDay$total.steps,na.rm=TRUE)


```

* Mean number of steps taken per day - ```r as.integer(meanStepsPerDay) ```        
* Median number of steps per day - ```r medianStepsPerDay```  


## What is the average daily activity pattern?

```{r}

stepsByTime<-group_by(select(activityData,interval,steps),interval)
stepsByTime<-summarise(stepsByTime,mean(steps,na.rm=TRUE))

names(stepsByTime)<-c("Interval","Average.Steps")

timeMin<-stepsByTime$Interval%%100
timeHour<-floor(stepsByTime$Interval/100)
stepsByTime$timeOfDay<-paste(timeHour,":",timeMin)

timeAtmax<-stepsByTime[stepsByTime$Average.Steps==max(stepsByTime$Average.Steps),]

g<-ggplot(stepsByTime,aes(y=Average.Steps,x=Interval,color=topo.colors(1)))
g<-g+geom_line()
g<-g+ theme(legend.position="none")
g<-g+labs(y="Average Steps",x = "Time in a day (5-min interval)")
g<-g+ggtitle(paste("Maximum number of steps reported at",timeAtmax$timeOfDay))
print(g)

```

At 5-minute interval, on average across all the days maximum number of steps 
reported at ```r timeAtmax$timeOfDay```

## Imputing missing values

```{r}
summary(activityData)
countNa<-sum(is.na(activityData$steps))


```

Total number of missing values ```r countNa```

**Strategy to remove NAs** Replace NAs with average steps taken in the 5-min 
time slot on other days

```{r}

activityMerged<-merge(activityData,stepsByTime,by.x="interval",by.y="Interval")

activityMerged<-transform(activityMerged,steps=ifelse(is.na(steps),Average.Steps,steps))

activityDataNaRemoved<-select(activityMerged,steps,date,interval)

summary(activityDataNaRemoved)

```

Redraw the histogram after replacing NAs

```{r, warning=FALSE}
activityByDay<-group_by(select(activityDataNaRemoved,date,steps),date)
stepsPerDay<-summarise(activityByDay,sum(steps))

names(stepsPerDay)[2]<-"total.steps"

g<-ggplot(stepsPerDay,aes(total.steps))
g<-g+geom_histogram(binwidth = 2500,col="white")
g<-g+ggtitle("Steps taken per day")
g<-g+labs(y="Frequency",x = "Total number of steps per day (NAs removed)")
print(g)


meanStepsPerDay<-mean(stepsPerDay$total.steps)
medianStepsPerDay<-median(stepsPerDay$total.steps)


```
After imputing missing values  
* Mean number of steps taken per day - ```r as.integer(meanStepsPerDay) ```  
* Median number of steps per day - ```r as.integer(medianStepsPerDay)```   

These numbers are not significantly different than before

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activityDataNaRemoved<-transform(activityDataNaRemoved,
                                 date=as.Date(date,"%Y-%m-%d"))

activityDataNaRemoved<-transform(activityDataNaRemoved, weekday=weekdays(date))

activityDataNaRemoved<-transform(activityDataNaRemoved, 
                                 wday=ifelse(weekday %in% c("Saturday","Sunday"),
                                             "Weekend","Weekday"))
                                   


stepsByTime<-group_by(select(activityDataNaRemoved,interval,steps,wday),
                      interval,wday)
stepsByTime<-summarise(stepsByTime,mean(steps))

names(stepsByTime)<-c("Interval","wday","Average.Steps")


g<-ggplot(stepsByTime,aes(y=Average.Steps,x=Interval))
g<-g+geom_line()
g<-g+facet_grid(wday~.)
g<-g+ theme(legend.position="none")
g<-g+labs(y="Average Steps",x = "Time in a day (5-min interval)")
g<-g+ggtitle("Difference between averge steps taken in Weekday vs Weekend")
print(g)

```
